---
title: "Analysis"
author: "Leeza Sergeeva"
date: "10/4/2020"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

## Setup

```{r set-parameters}
# choose the state to manipulate
state_to_analyze <- "Florida"

# sanitizing the name of the state for output files
state_no_spaces <- gsub(state_to_analyze,
                        pattern = " ",
                        replacement = "_")

# set global data file name to use
raw_file_to_process <- "Data/Raw/applemobilitytrends-2020-10-03.csv"

# get base name for the output file without extention
base_name <- tools::file_path_sans_ext(basename(raw_file_to_process))

```

```{r load-packages-functions}
# load necessary packages
library("dplyr")
library("readr")
library("ggplot2")
library("tidyr")
library("lubridate")
library("devtools")

# load necessary functions
source("Code/Functions/state_mobility_data.R")
source("Code/Functions/count_counties_and_cities_traffic.R")
source("Code/Functions/create_long_pivot_table.R")
```

## Project Description

This markdown file will use `ggplot2` [@ggplot2], `dplyr` [@dplyr], and `readr` [@readr] packages to analyze the Apple maps mobility data since the beginning of the COVID-19 pandemic (initial shutdown) and how mobility changed throughout the year..

The goal is to create a subset to a particular state, tally up the number of cities and counties in that state with mobility data, and then make several plots to visualize those data subsets.

The state that this document will analyze is `r state_to_analyze`.

## Subsetting Data

```{r state_subset}
# subset out to state selected above in the document
state_subset <- state_mobility_data(
  input_file_name = raw_file_to_process,
  state = state_to_analyze
)
```

```{r state_subset_long}
# convert subset state data to long format pivoting dates
long_subset <- create_long_pivot_table(
  input_file_name = paste0("Output/State_Data_Wide/",
                           base_name,
                           "_",
                           state_no_spaces,
                           ".csv")
  )
```
```{r state_subset_tally}
# tally up the numbers of cities in the state with mobility data
subset_tally <- city_county_mobility_data(
  input_state_file = paste0("Output/State_Data_Wide/",
                           base_name,
                           "_",
                           state_no_spaces,
                           ".csv")
  )
```

## Figures

```{r plot-tally-data}
# create grouped bar plot of the tallied data to include in the knitted
# document, and save that figure in output folder as .png file
plot_tally <- ggplot(data = subset_tally,
                     aes(x = geo_type,
                         y = n,
                         fill = transportation_type)) +
  geom_col(position = position_dodge()) +
  labs(title = paste("Number of cities and/or counties in",
                      state_to_analyze,
                      "with mobility data"),
       x = "Type of area",
       y = "Number of areas")

ggsave(plot = plot_tally,
       filename = paste0("Output/Figures/",
                         base_name,
                         "_",
                         state_no_spaces,
                         "_tally_plot.png"))

plot_tally
```

```{r plot-time-series}
# make a line plot of relative driving mobility across the state with date
# on the x axis and relative mobility on the y axis
timeseries_plot <- long_subset %>%
  filter(transportation_type == "driving") %>%
  group_by(date) %>%
  summarize(mean_mobility = mean(rel_mobility)) %>%
  ggplot(aes(x = lubridate::ymd(date),
             y = mean_mobility)) +
  geom_line() +
  labs(title = paste("Statewide mean relative mobility driving leevls in",
                     state_to_analyze,
                     "during COVID"),
       x = "Date",
       y = "Mean relative mobility")

ggsave(plot = timeseries_plot,
       filename = paste0("Output/Figures/",
                         base_name,
                         "_",
                         state_no_spaces,
                         "_timeseries_plot.png"))
timeseries_plot
```

## Session Info

```{r session-info}
devtools::session_info()
```

## Sources Cited
